% !TeX spellcheck = en_US
% !TeX root = tikz-ext-manual.tex
% Copyright 2022 by Qrrbrbirlbel
%
% This file may be distributed and/or modified
%
% 1. under the LaTeX Project Public License and/or
% 2. under the GNU Free Documentation License.
%
\clearpage
\section{Positioning Plus}
\label{library:positioning-plus}
\tikzsetnextfilename{posplus}
\begin{tikzlibrary}{ext.positioning-plus}
  With the help of the \referenceLibraryandIndexO{positioning} and the \referenceLibraryandIndexO{fit} library this extends the placement of nodes.
\end{tikzlibrary}

\subsection{Useful corner anchors}
The anchors |corner north east|, |corner north west|, |corner south west| and |corner south east|
are defined as ``generic anchors'', i.\,e. they are defined for all shapes.
This is mostly useful for the placement of circular shapes.
\begin{stylekey}{/tikz/corner above left=\opt{\meta{specification}} (default 0pt)}
  Similar as \referenceKeyandIndexO{above left} of the \tikzname\space library |positioning|
  but uses the |corner north west| anchor.
\end{stylekey}
\begin{stylekey}{/tikz/corner below left=\opt{\meta{specification}} (default 0pt)}
  Similar as \referenceKeyandIndexO{below left} of the \tikzname\space library |positioning|
  but uses the |corner south west| anchor.
\end{stylekey}
\begin{stylekey}{/tikz/corner above right=\opt{\meta{specification}} (default 0pt)}
  Similar as \referenceKeyandIndexO{above right} of the \tikzname\space library |positioning|
  but uses the |corner north east| anchor.
\end{stylekey}
\begin{stylekey}{/tikz/corner below right=\opt{\meta{specification}} (default 0pt)}
  Similar as \referenceKeyandIndexO{below right} of the \tikzname\space library |positioning|
  but uses the |corner south east| anchor.
\end{stylekey}

\begin{codeexample}[width=.6\linewidth,preamble=\usetikzlibrary{ext.positioning-plus}]
\Huge
\begin{tikzpicture}
\node[name=s,shape=circle,shape example]
  {Circle\vrule width 1pt height 2cm};
\foreach \anchor/\placement in {
  north west/above left, north/above, north east/above right,
  west/left, center/above, east/right,
  mid west/right, mid/above, mid east/left,
  base west/left, base/below, base east/right,
  south west/below left, south/below, south east/below right,
  text/left, 10/right, 130/above}
  \draw[shift=(s.\anchor)] plot[mark=x] coordinates{(0,0)}
    node[\placement] {\scriptsize\texttt{(s.\anchor)}};
\draw (s.corner north west) rectangle (s.corner south east);
\foreach \anchor/\placement in {
  corner north west/above left, corner north east/above right,
  corner south west/below left, corner south east/below right}
  \draw[red,shift=(s.\anchor)] plot[mark=x] coordinates{(0,0)}
    node[\placement] {\scriptsize\texttt{(s.\anchor)}};
\end{tikzpicture}
\end{codeexample}

\subsection{Useful placement keys for vertical and horizontal alignment}
\begin{stylekey}{/tikz/north left=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{left} but aligns the nodes at their |north| border.
  
  This is basically the same as |left=of reference.north west, anchor=north east|.
  %
\begin{codeexample}[preamble=\usetikzlibrary{ext.positioning-plus}]
\begin{tikzpicture}[nodes=draw]
\node[minimum height=2cm] (a) {};
\node[minimum height=3cm, north right=of a] {};
\end{tikzpicture}
\end{codeexample}
\end{stylekey}
\begin{stylekey}{/tikz/north right=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{right} but aligns the nodes at their |north| border.
  
  This is basically the same as |left=of reference.north east, anchor=north west|.
\end{stylekey}
\begin{stylekey}{/tikz/south left=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{left} but aligns the nodes at their |south| border.
  
  This is basically the same as |left=of reference.south west, anchor=south east|.
\end{stylekey}
\begin{stylekey}{/tikz/south right=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{right} but aligns the nodes at their |south| border.
  
  This is basically the same as |left=of reference.south east, anchor=south west|.
\end{stylekey}
\begin{stylekey}{/tikz/west above=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{above} but aligns the nodes at their |west| border.
  
  This is basically the same as |left=of reference.north west, anchor=south west|.
\end{stylekey}
\begin{stylekey}{/tikz/west below=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{below} but aligns the nodes at their |west| border.
  
  This is basically the same as |left=of reference.south west, anchor=north west|.
\end{stylekey}
\begin{stylekey}{/tikz/east above=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{above} but aligns the nodes at their |east| border.
  
  This is basically the same as |left=of reference.north east, anchor=south east|.
\end{stylekey}
\begin{stylekey}{/tikz/east below=\opt{\meta{specification}} (default 0pt)}
  Like \referenceKeyandIndexO{below} but aligns the nodes at their |east| border.
  
  This is basically the same as |left=of reference.south east, anchor=north east|.
\end{stylekey}

The same exist for the recently introduces corner anchors, too.
\begin{stylekey}{/tikz/corner north left=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/north left| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner north right=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/north right| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner south left=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/south left| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner south right=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/south right| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner west above=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/west above| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner west below=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/west below| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner east above=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/east above| but uses the new corner anchors.
\end{stylekey}
\begin{stylekey}{/tikz/corner east below=\opt{\meta{specification}} (default 0pt)}
  The same as |/tikz/east below| but uses the new corner anchors.
\end{stylekey}

While the \meta{specification} of all these keys still accept the same form as with \tikzname,
the |ext.positioning-plus| library extends this even more.

The specification after |of| can contain a list of coordinates
(like the |fit| key of the |fit| library).
This means that the new node will be placed in relation to a rectangular bounding box
that fits around all this nodes in the list.

If this list is prefixed with \verb!|!, |-| or |+|,
the new node will also have the same height (\verb!|!),
the same width (|-|) or both as this bounding box.

\begin{codeexample}[preamble=\usetikzlibrary{ext.positioning-plus}]
\begin{tikzpicture}[nodes=draw]
\node                     (A) {A};
\node[below=of A]         (BCD) {BCD};
\node[right=of |(A)(BCD)] (c) {};
\node[below=.5:of -(A)(BCD)] (d) {};
\draw[help lines] (BCD.south west) -- (c.south east)
                  (BCD.north east) -- (d.south east);
\end{tikzpicture}
\end{codeexample}

This functionality is also available without the placement:
\begin{stylekey}{/tikz/fit bounding box=\meta{list of coordinates}}
  Creates a rectangular node with the name |fit bounding box| that encompasses
  the \meta{list of coordinates}.
\end{stylekey}
\begin{stylekey}{/tikz/span vertical=\meta{list of coordinates}}
  Creates a rectangular node with the name |fit bounding box| that encompasses
  the \meta{list of coordinates} \emph{and} sets the \referenceKeyandIndexO[/pgf/]{minimum height} to
  the height of this bounding box.
\end{stylekey}
\begin{stylekey}{/tikz/span horizontal=\meta{list of coordinates}}
  Creates a rectangular node with the name |fit bounding box| that encompasses
  the \meta{list of coordinates} \emph{and} sets the \referenceKeyandIndexO[/pgf/]{minimum width} to
  the width of this bounding box.
\end{stylekey}
\begin{stylekey}{/tikz/span=\meta{list of coordinates}}
  Is a combination of |/tikz/span vertical| and |/tikz/span horizontal|.
\end{stylekey}

As you maybe noticed in the example above, the \meta{specification} also allows a prefix
delimited by |:| which the |node distance| will be multiplied to with for the placement.%
\footnote{This is probably more useful when \referenceKeyandIndexO{on grid} is used.}

\endinput