% !TeX spellcheck = en_US
% !TeX root = tikz-ext-manual.tex
% Copyright 2022 by Qrrbrbirlbel
%
% This file may be distributed and/or modified
%
% 1. under the LaTeX Project Public License and/or
% 2. under the GNU Free Documentation License.
%

\section{Shape: Uncentered Rectangle}
\begin{purepgflibrary}{ext.shapes.uncenteredrectangle}
  A rectangle that has a variable horizontal center with three node parts.
  \inspiration{UncRectCD-Q,UncRectForest-Q}{UncRectCD-A,UncRectForest-A}
\end{purepgflibrary}
\begin{ext_shape}{uncentered rectangle}

For some alignment problems, this shape could be useful.

It has three node parts: the standard |text| part,
the |left| part that is to the left of |text|
and the |right| part that is to the right of |text|.

When edges are to be connected with this shape, the
following key changes to which inner center this shape will
calculate the appropriate point on the border.
\begin{key}{/\pgfext/uncentered rectangle center=\meta{left}\textrm{ or }\meta{text}\textrm{ or }\meta{right}\textrm{ or }\meta{real} (initially text)}\keycompat{pgf}
  Sets the center that is to be used for connecting edges.
  
  This will also move the anchors |north|, |mid|, |base| and |south| along.
  In the picture below, this are marked red.
\end{key}

\begin{key}{/\pgfext/uncentered rectangle use saved center=\meta{true}\textrm{ or }\meta{false} (default true)}\keycompat{pgf}
When this is set to true, the border anchors will use the horizontal center that was used when
the node was created.
\end{key}

For support of the \referenceLibraryandIndexO{cd} library of the |tikz-cd| package,
this shape also supports a dynamic $y$ value for its anchors |center|, |west| and |east|.
\begin{key}{/\pgfext/uncentered rectangle center yshift=\meta{dimension} (initially \{\})}\keycompat{pgf}
  This determines the distance between the baseline and the |center| anchors.
  
  If \meta{dimension} is empty, the real vertical center will be used.
  
  For use with |cd|, set this to |axis_height|.
\end{key}
%\pagebreak
\begin{codeexample}[preamble=\usepgflibrary{ext.shapes.uncenteredrectangle}]
\begin{tikzpicture}[style north/.style=red, style south/.style=red, style center/.style=red, style base/.style=red, style mid/.style=red]
\Huge
\node[shape example, name=n, ext_uncentered rectangle]
  {centered \nodepart{left} Un \nodepart{right} \space Rectangle\vrule width 1pt height 2cm}
  foreach \anchor/\pos in {
   north west/above left, north/below, north east/above right, real north/above,  left north/above, right north/above, text north/above,
         west/left,      center/above,       east/right,       real center/above, left center/above,right center/above,text center/below,
     mid west/left,         mid/left,    mid east/right,       real mid/above,    left mid/above,   right mid/above,   text mid/above,
    base west/left,        base/right,  base east/right,       real base/below,   left base/below,  right base/below,  text base/below,
   south west/below left, south/above, south east/below right, real south/below,  left south/below, right south/below, text south/below,
                             10/right,        130/below,                          left/left,        right/right,       text/right}{
    plot[mark=x, only marks] coordinates {(n.\anchor)}
    node[inner sep=.1em, style \anchor/.try, style/.expand once=\pos] {\tiny\ttfamily\anchor}};
\end{tikzpicture}
\end{codeexample}
\end{ext_shape}

\begin{tikzlibrary}{ext.shapes.uncenteredrectangle}
This library extends the \referenceLibraryandIndexO{cd} library (from the |tikz-cd| package)
so that it can be used with the |uncentered rectangle| shape.

\inspirationQ{UncRectCD2-Q}
\end{tikzlibrary}

This library provides only one key.
\begin{stylekey}{/\tikzext/tikz-cd fix}\keycompat{tikz-ext}
This key installs various \enquote{fixes} to the \referenceKeyandIndexO[/tikz/commutative diagrams/]{every diagram} style:

\begin{itemize}
\item Firstly, is defines a \referenceKeyandIndexO{matrix of math nodes} key (only for the \referenceEnvironmentandIndexO{tikzcd} environment)
      which allows to toggle the \referenceKeyandIndexO[/tikz/commutative diagrams/]{math mode} for each node.%
      \footnote{Due to a bug with \referenceKeyandIndexO{execute at end node}, the \enquote{automatic} math mode in matrices can't be used
        with multipart nodes.}
\item The helpful macro |\uncrec| will be installed.
\begin{command}{\uncrec\marg{left}\marg{center}\marg{right}}
  When used as the content of a |ext_uncentered rectangle| shape,
  the node parts will be setup so that \meta{left} is in the left part of the node part etc.
\end{command}
\item Since math mode will be disabled with the |ext_uncentered rectangle|, it is automatically enabled for each node part with |\uncrec| but it can be disabled with the following key.
\begin{key}{/tikz/uncrec math mode=\meta{true}\textrm{ or }\meta{false} (default true)}
When enabled the contents of |\uncrec| will be set in math mode.
\end{key}
\item For easy access to the |uncentered rectangle| shape, the following keys are available inside a Commutative Diagram.
\begin{stylekey}{/tikz/uncrec=\meta{left}\textrm{ or }\meta{text}\textrm{ or }\meta{right}\textrm{ or }\meta{real} (initially text)}
This key sets the shape to |ext_uncentered rectangle| and \referenceKeyandIndex[/\pgfext/]{uncentered rectangle center} to its argument.
\end{stylekey}
\begin{stylekey}{/tikz/commutative diagrams/install uncentered rectangle in column=\meta{column}}
All nodes in column \meta{column} will be set to the |ext_uncentered rectangle| shape.
\end{stylekey}
\end{itemize}
\end{stylekey}

\begingroup
\tikzexternaldisable
%\catcode`\|=12
\begin{codeexample}[leave comments, width=8cm, preamble=\usetikzlibrary{cd, ext.shapes.uncenteredrectangle}]
\tikzextset{tikz-cd fix}
\newcommand*\C[1]{C_{\%_{#1}}}
\begin{tikzcd}[
  sep=tiny,
  arrows={-, gray},
  cells={font=\strut, inner xsep=.2ex, inner ysep=.1ex},
  install uncentered rectangle in column=3
]
\C{1} \drar &         & \uncrec{}{m_{r_1}}{{} = \C{2}-C_\%} \dlar\\
            & C_\% \\
\C{2} \urar &         & \uncrec{}{m_{r_2}}{{} = \C{1}-C_\%} \ular
\end{tikzcd}
\end{codeexample}
\begin{codeexample}[leave comments, width=8cm, preamble=\usetikzlibrary{cd, ext.shapes.uncenteredrectangle}]
\tikzextset{tikz-cd fix}
\begin{tikzcd}[install uncentered rectangle in column/.list={1,2}]
  \uncrec{S \supset {}}{U_\tau}{}                                      \arrow[r, "\varphi_0"]
                                                                       \arrow[d, "\tau", "\sim"']
& \uncrec{}{U_\pi}{{} \subset T}                                       \arrow[d, "\pi",  "\sim"']
\\
  \uncrec{\operatorname{Bl}_{(0,0)}(\mathbb{A}^2) \supset{}}{V_\tau}{} \arrow[r, "\epsilon"]
& \uncrec{}{V_\pi}{{} \subset \mathbb{A}^2}
\end{tikzcd}
\end{codeexample}
\endgroup
\endinput